A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-06-09 bioc-workflow.Rmd
TODO: other plot2D methods
replicates section and example

```{r env, echo=FALSE}
library("BiocStyle")
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRoloc"))
suppressPackageStartupMessages(library("pRolocdata"))
```

@@

## Replication

See issue #6.

@@

@@

@@

```{r visualiseOpt, eval=FALSE}
(best <- getParams(params))
plot(params)
levelPlot(params)
```
```{r visualiseOptHide, include=FALSE}
grid.arrange(plot(params), levelPlot(params), ncol = 2)
```

>>>>

```{r env, echo=FALSE}
library("BiocStyle")
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRoloc"))
suppressPackageStartupMessages(library("pRolocdata"))
```

@@

When more than 2 data are to be combine and/or too many proteins were
not consistently assayed, leading to too many proteins being filtered
out, we sould suggest to implement an ensemble of classifiers voting
on protein-sub-cellular niche membership over the output of several
experiments (see section *Supervised machine learning* for the
description of sub-cellular assingments).

@@

**TODO** Mention that there are other dimensionality reduction
techniques. Advantages, disadvantages? See
[this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
very interesting gist with intriguing tSNE examples.

@@

# Replication

With the aim of maximising the sub-cellular resolution and,
consequently, the reliability in protein sub-cellular assignments, we
follow the advice in [@Trotter:2010] and combine replicated spatial
proteomics experiments and described above.

Trotter et al. have shown a significant improvement in
protein–organelle association upon direct combination of single
experiments, in particular when these resolve different subcellular
niches.

Direct comparisons of individual fractions in replicated experiments
does not provide an good, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and fraction
collection, the quantitative channels (loosely referred to as
fractions) do not correspond to identical fractions along the
gradient. As we see in the table below (taken from the `hyperLOPIT`'s
`pData`), different sets of fractions are combined to obtain enough
material and optimise acurate quantitation.

```{r pdtab, echo=FALSE}
i <- hyperLOPIT$Fraction.No %in% 7:10
kable(pData(hyperLOPIT)[i, -3], row.names = FALSE)
```

On the figure below, we compare the relative intensities of fractions
7 to 10 on scatter plots, highlighting different marker sets, in
particular mitochondrion and chromatin. These differences result from
different peak fractions for these niches.

```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12}
sel0 <- fData(hyperLOPIT)$markers != "unknown"
cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
cls[!sel0] <- getUnknowncol()

par(mfrow = c(2, 4))
for (i in c(7, 8, 10, 9)) {
    plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
         main = paste("Fraction", pData(hyperLOPIT)[i, "Fraction.No"]))
    points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
    grid()
    abline(0, 1, col = "red")
}
plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "n")
addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)

i <- fData(hyperLOPIT)$markers == "Mitochondrion"
plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
         featureNames(hyperLOPIT)[i], type = "l",
         mcol = "orange")
title(main = "Mitochondrion")
## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
legend("topleft", paste("Replicate", 1:2),
       col = c("#F781BF", "orange"),
       lty = 1, bty = "n")

i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
         featureNames(hyperLOPIT)[i], type = "l")
## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
title(main = "Chromatin")
legend("topleft", paste("Replicate", 1:2),
       col = c("#984EA3", "steelblue"),
       lty = 1, bty = "n")
```

The relevant comparison unit is not a single fraction, but rather
complete protein occupancy profiles, which are best visualised as a
whole on a PCA plot. As such, we prefer to focus on the direct
comparison of individual replicate PCA plots, assuring that each
displays acceptable sub-cellular resolution. Note that in the code
chunk below, we mirror the y-axis to represent the two figures with
the same orientation.

```{r plot2Drep, fig.width = 12}
par(mfrow = c(1, 2))
plot2D(hl[hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
plot2D(hl[hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
       mirrorY = TRUE)
```

@@

```{r visualiseOpt, eval=FALSE}
(best <- getParams(params))
plot(params)
levelPlot(params)
```
```{r visualiseOptHide, echo=FALSE, fig.width = 12}
grid.arrange(plot(params), levelPlot(params), ncol = 2)
```
